At the beginning of 2018 the Open Data Lab was an idea, a sketch on a whiteboard. Today the Open Data Lab has over 100 users and students at the University of Virginia have used it to conduct research published by IEEE. In just over a year a small team of dedicated individuals from the Data Science Institute at UVA produced a product that furthered the educational and research mission of the University. Today students at UVA can do things they could not do before this project launched.

I am extremely honored to have been asked to lead this effort. The work done by our collaboration is of the very best calibre and I am proud of the work we have done and will do in the future. As project manager I feel I have the easiest job on the team. I ask everyone for ideas and they put forth not just an enormous number of ideas but also brilliant ideas. The hardest part for me is to decide which ideas will get us to where we want to go. Without the team Phil Bourne has built my job would be impossible. With this team I do not see any limit to what we can achieve.

The first year of the Open Data Lab was one of exploration. We discovered new tools, in particular the value of Project Jupyter. We also examined our potential users and developed archetypes which we can use to guide what we build. Putting all of this together we are currently developing a strategic plan for the next three years. I do not have anything certain to say yet, so stay tuned for more news.

\begin{flushright}

\bigskip

Open Data Lab Project Manager

\bigskip

Peter Alonzi

\end{flushright}
